---
title: "Project Practical Machine Learning jdemoya"
author: "Jose De Moya"
date: "Saturday, October 07, 2014"
output: html_document
---

The goal of the project is predict the manner in with the people in the files did the exercises.

That is the procedure:

First I load the data in two variables:

```{r}
library(caret);
setwd("D:/Cursos/Coursera/Data Science/Practical Machine Learning");
training <- read.csv("pml-training.csv"); 
testing <- read.csv("pml-testing.csv");
dim(training);
dim(testing);
```

```{r, eval=FALSE}
summary(cars)
```

The variables that are very correlated that must be excluded of the model are obtained from but first We must do a preprocess because the factor variables don't work with correlations 

```{r}
str(training);
```

First I remove the variables with zero variance because they are not good predictors

```{r}
nsv <- nearZeroVar(training,saveMetrics=TRUE);
nsv
nsv[nsv$nzv == TRUE,];
```

The factor variables can be converted into numeric or dates when apply


```{r}
n <- length(names(training));
m <- length(training[,1]);
training1 <- training;

for (i in 1:n){
  if (!is.factor(training1[,i])){
    training1[is.na(training1[,i]),i] <- 0;
  }  
  else{
    x <- c(x,i);
  }
}

for (i in 1:n){
  if (is.factor(training1[,i])){
    f <- training1[,i];
    training1[,i] <- as.character(levels(f))[f];
  }
}
mcor <- cov2cor(cov(training1[,-x]));
mcor>=0.7;
```
